---
title: "A joint model for the initiation and cessation of electronic cigarettes and conventional cigarettes with time-dependent covariates in a longitudinal study"
author: Zhi Yang PhD, Kiros Berhane PhD, Adam Leventhal PhD, Mengyu Liu MS, Jessica L. Barrington-Trimis PhD1, Duncan Thomas PhD 
date: "`r Sys.Date()`"
output:
  rmdformats::html_clean:
    thumbnails: false
    highlight: kate
---


```{r setup, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)
library(ggplot2)

## Global options
options(max.print="100")
opts_chunk$set(echo=TRUE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=FALSE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

# Introduction 
This vignette is provided to run the models described in the paper "A joint polytomous logistic regression model for the initiation and cessation of electronic cigarettes and conventional cigarettes with time-dependent covariates in a longitudinal study". 

# Methods 
All functions are stored in `model_functions.R` which includes key functions like `Initiate`, `SimulateData` and `Analyze`. 
## Simulating data 
```{r}
source("model_functions.R")
# Parameter specifications 
nT <- 5 # number of time points (including the baseline)
V <- 4 # number of covariates
N <- 1000

simstat <- Initiate(nT=nT, V=V)
set.seed(2020)
sData <- SimulateData(param=simstat, N=N)

X <- sData[["X"]]
Y <- sData[["Y"]]
```

## Data overview 
`simstat` is a list containing parameters like $\lambda$s, $\mu$s, $\alpha$s, $\eta$s, $\beta$s, $\gamma$s, and $\delta$s. 
```{r}
simstat
```

`X` is a list of `N` matrices of nT $\times$ V, each of which contains `V` covariates over `nT` time points for one observation. 
```{r}
head(X, n = 2)
```

`Y` is a list of `N` matrices of nT $\times$ V, each of which contains two outcomes (0: No; 1: Yes) over `nT` time points for one observation. 
```{r}
head(Y, n = 2)
```

## Running the joint polytomous logistic regression models
```{r cache=TRUE}
res_joint <- Analyze(simstat, X, Y)
```

# Results
The model yields the coefficients and standard errors of all parameters listed by `simstat`. 
```{r}
# coefficients 
coef_joint <- vec2list(res_joint$par)
# standard errors 
sigma_joint <- vec2list(sqrt(diag(solve(-res_joint$hessian))))
# create 95% CI plots 

head(coef_joint, n=2)
head(sigma_joint, n=2)
```

We displayed the odds ratio plots for important variables $\beta$s and $\gamma$s. 
```{r oddsratio, fig.width=8}
df <- data.frame(type = factor(
                   rep(c("Initiation", "Cessation"), each = 2), 
                   levels = c("Initiation", "Cessation")
                 ),
                 cig = c("E-cigarettes", "C-cigarettes", 
                         "E-cigarettes", "C-cigarettes"),
                 mean = c(coef_joint$beta,
                          coef_joint$gamma), 
                 sd = c(sigma_joint$beta,
                        sigma_joint$gamma))

df$lb <- exp(df$mean + qnorm(0.025)*df$sd)
df$ub <- exp(df$mean + qnorm(0.975)*df$sd)
df$mean <- exp(df$mean)

ggplot(df, aes(x = cig, y = mean)) +
  geom_point(position=position_dodge(width=0.3), size=2) +
  geom_hline(yintercept = 1.0, linetype="dotted", size=1) + 
  scale_y_continuous(trans='log2', 
                     breaks = c(round(df$lb, 1), 
                                round(df$ub, 1), 1)) +
  scale_color_manual(values = c("grey", "black")) +
  guides(color = guide_legend(reverse = TRUE)) +
  geom_errorbar(aes(ymin = lb, ymax = ub),
                position = position_dodge(0.3), width = 0.25) +
  facet_grid(type~.)+
  coord_flip() +
  ylab("Odds ratio") +
  theme_bw()+
  theme(text = element_text(size=15),
        axis.text.x = element_text(size=8),
        panel.grid.minor = element_blank(),
        legend.position = "top", 
        legend.title = element_blank(), 
        axis.title.y = element_blank())
```


